DINO Methods for Style Transfer

### 2022-04-19 paper/Paper.tex
another draft

\begin{abstract}
We combine DINO \cite{DINO}, a self-supervised learning method that captures feature representations of input images, with StyTr2 \cite{ImageStyleTransformer}, a style transfer transformer that renders images with artistic style, and explore the nature and robustness of the resulting model.
\end{abstract}

\section{Introduction}
Style transfer is the problem of changing an image's artistic style while maintaining its ``content'', namely the object it depicts. Conventional models use a CNN to extract features from an image, take in another image indicating the desired artistic style, and attempt to apply that style to the features of the original input image. However, it has been shown that due to the locality resulting from shared weights, CNN based models struggle to capture global information of input images and suffer from feature losses \cite{ImageStyleTransformer}. To resolve these issues, models that use transformers instead of CNNs for feature and style extraction have been proposed and show better results \cite{ImageStyleTransformer}. The model StyTr2 proposed in \cite{ImageStyleTransformer} takes in a content image and a style image, uses two transformer encoders to capture the content and style, then renders the content with that style using a thrid transformer decoder. In our project, we will substitute the DINO method and models from \cite{DINO} for the encoder components and explore the effects on the task of style transfer. The DINO method is good at extracting content even if the input images are distorted or partial, so in particular, we will implement the models and conduct experiments to explore the following questions:
\begin{enumerate}
	\item By using DINO to extract content, can we obtain a style transfer that is more robust to distortion of the content image?
	\item Does DINO's ability to capture an image's global property from local pieces give it more robustness as a style encoder?
	\item Are the results of this new style transfer method better than previous methods?
\end{enumerate}

@@

\cite{DINO} came up with a new self-supervised learning method, named DINO (for Knowledge Distillation with No Labels) that aims to create better feature representations from its inputs by adopting a teacher-student paradigm. The student gets a difficult version (lots of augmentation) of the image while the teacher gets an easy version (minimal augmentation), and then the models are trained to aim for the same feature representation through cross entropy loss. Their experimental results found that the vision transformer's class token contained explicit information about the segmentation of objects in images, while also performing as very accurate classifiers. In this project, we aim to apply the better feature representation that the DINO method yields (over supervised vision transformers), to the domain of image style transfer.

Self-supervised learning is a common technique used in style transfer, due to the obvious lack of target images available for a task of this nature. It becomes necessary to find ways for the model to supervise its own outfputs. Usual methods \cite{CNNStyleTransfer} involve checking the outputs of VGG for consistency in style and context. \cite{ImageStyleTransformer} proposes using vision transformers for this task, using two transformer encoders (one for content, one for style), and one transformer decoder. Transformers have been a hot rising topic in computer vision, and they claim that their method of using transformers for style transfer results in a decrease in content leak, as well as achieves unbiased stylization.

It becomes integral for there to be good feature representations of the content and style to construct a good style-transferred image. We believe that making use of pretrained DINO ViT models, as well as using the DINO method itself during training, can help achieve this.

@@

\subsection{DINO Training for Encoders}

To translate the success of DINO training on classification to style transfer we modify our algorithm and architecture. Now instead of simple a style encoder, there is a teacher style encoder $SE_t$ and a student style encoder $SE_s$, both using the architecture from \cite{ImageStyleTransformer}. This forces the students to learn distillations of the original images. Because style is a global property of an image, we applied this only to the style encoder and left the content encoder $CE$ as in \cite{ImageStyleTransformer}.

To train this model, as in the \cite{DINO}, from the original style image we generate two slightly modified global views of the image $(I_s)^g_1$ and $(I_s)^g_2$ as well as a collection more heavily modified local views, with the set of all views denoted $V$. The student is passed all the views whereas the teacher is only passed the global views. We then calculate 
\begin{align}
	\mathcal L_d &= \sum_{I \in \{(I_s)^g_1, (I_s)^g_2\}} \sum_{I' \in V - \{ I \}} H(SE_t(I), SE_s(I'))
\end{align} 
Where $H(a,b) = -a \log (b)$. This loss is what forces our student style encoder to learn these distilled representations. We can average $SE_s((I_s)^g_1)$ and $SE_s(I_s)^g_2$ to get the encoded style, and we can take $CE(I_c)$ to get the content. From this we can generate $I_o$, the output image, using the decoder $D$. We can also pass the style to both inputs of $D$ to get $I_{ss}$ and similarly we get $I_{cc}$. Now we can use $\mathcal L_c$, $\mathcal L_s$, $\mathcal L_{id1}$,$\mathcal L_{id2}$ from \cite{ImageStyleTransformer}. We can then combine these into the final loss
\begin{align}
	\mathcal L &= \lambda_d \mathcal L_d + \lambda_c \mathcal L_c + \lambda_s \mathcal L_s + \lambda_{id1} \mathcal L_{id1} + \lambda_{id2} \mathcal L_{id2}
\end{align}
where the $\lambda$'s are weightings. We call this model D).

We also define model E) where we modify (1) to
\begin{align}
	\mathcal L_d &= \sum_{I \in \{(I_s)^g_1, (I_s)^g_2\}} \sum_{I' \in V - \{ I \}} H(D(SE_t(I), CE(I_c)), D(SE_s(I'), CE(I_c))).
\end{align}
Where calculating the loss after the decoder should weight the most relevant student and teacher differences higher.

@@

\includegraphics[scale=0.35]{comparison}

@@

\subsection{Comparisons with Other Methods}

\subsubsection{Timing Information}

\subsubsection{Qualitative Evaluation}

\subsubsection{Quantitative Evaluation}

>>>>

\begin{abstract}
We apply DINO \cite{DINO}, a self-supervised learning method that captures feature representations of input images, to style transfer by extending StyTr2 \cite{ImageStyleTransformer}, a style transfer transformer that renders images with artistic style, and explore the nature and robustness of the resulting model. We explore using the vision transformers pretrained with the DINO method to encode image style, and we also explore apply the DINO training methods directly to StyTr2. We discuss the relative qualitative and quantitative performance of our new models compared to the state of the art.
\end{abstract}

\section{Introduction}
Style transfer is the problem of changing an image's artistic style while maintaining its ``content'', namely the object it depicts. Conventional models use a CNN to extract features from an image, take in another image indicating the desired artistic style, and attempt to apply that style to the features of the original input image. However, it has been shown that due to the locality resulting from shared weights, CNN based models struggle to capture the global information of input images and suffer from feature losses \cite{ImageStyleTransformer}. To resolve these issues, models that use transformers instead of CNNs for feature and style extraction have been proposed and show better results, like StyTr2 \cite{ImageStyleTransformer}. This model takes in a content image and a style image, then uses two transformer encoders to capture the content and style, then renders the content with that style using a transformer decoder. In our project, we will substitute the DINO method and models from \cite{DINO} for the encoder components and explore the effects on style transfer. The DINO method is good at extracting content even if the input images are distorted or partial, so in particular, we will implement the models and conduct experiments to explore the following questions:
\begin{enumerate}
	\item Does DINO's ability to capture an image's global property from local pieces give it more robustness as a style encoder?
	\item Are the results of this new style transfer method better than previous methods?
\end{enumerate}

@@

\cite{DINO} came up with a new self-supervised learning method, named DINO (for Knowledge Distillation with No Labels) that aims to create better feature representations from its inputs by adopting a teacher-student paradigm. The student gets multiple difficult views (lots of augmentation), called the local views of the image while the teacher gets an easy view (minimal augmentation), called the global view. Then the student is trained to aim for the same feature representation through cross entropy loss. Their experimental results found that the vision transformer's class token contained explicit information about the segmentation of objects in images, while also performing as very accurate classifiers. In this project, we aim to apply the better feature representation that the DINO method yields (over supervised vision transformers), to the domain of image style transfer.

Self-supervised learning is a common technique used in style transfer, due to the obvious lack of target images available for a task of this nature. It becomes necessary to find ways for the model to supervise its own outputs. Usual methods \cite{CNNStyleTransfer} involve checking the outputs of VGG for consistency in style and context. \cite{ImageStyleTransformer} proposes using vision transformers for this task, using two transformer encoders (one for content, one for style), and one transformer decoder. Transformers have been a hot rising topic in computer vision, and they claim that their method of using transformers for style transfer results in decreased content leak, as well as achieving unbiased stylization.

It is integral for there to be good feature representations of the content and style to construct a good style-transferred image. We believe that making use of pretrained DINO ViT models, as well as using the DINO method itself during training, can help achieve this.

@@

\subsection{DINO Training the Style Encoder}

To translate the success of DINO training on classification to style transfer we modify the algorithm and architecture of StyTr2. Now instead of simple transformer for the style encoder, there is a teacher style encoder $SE_t$ and a student style encoder $SE_s$, both using the architecture from \cite{ImageStyleTransformer}. We use the extra encoder to force the student to learn distillations of the original images. Because style is a global property of an image, we applied this only to the style encoder and left the content encoder $CE$ as it was in \cite{ImageStyleTransformer}.

To train this model, as in \cite{DINO}, from the original style image we generate two slightly modified global views of the image $(I_s)^g_1$ and $(I_s)^g_2$ as well as a collection more heavily modified local views, with the set of all views denoted $V$. The student is passed all the views whereas the teacher is only passed the global views. We then calculate 
\begin{align}
	\mathcal L_d &= \sum_{I \in \{(I_s)^g_1, (I_s)^g_2\}} \sum_{I' \in V - \{ I \}} H(SE_t(I), SE_s(I'))
\end{align} 
Where $H(a,b) = -a \log (b)$. This loss is what forces our student style encoder to learn these distilled representations. We can average $SE_s((I_s)^g_1)$ and $SE_s(I_s)^g_2$ to get the encoded style, and we can take $CE(I_c)$ to get the content. From this we can generate $I_o$, the output image, using the decoder $D$. We can also pass the encoded style to both inputs of $D$ to get $I_{ss}$ and similarly with the encoded content to get $I_{cc}$. Now we can use $\mathcal L_c$, $\mathcal L_s$, $\mathcal L_{id1}$,$\mathcal L_{id2}$ from \cite{ImageStyleTransformer}. We can then combine these into the final loss
\begin{align}
	\mathcal L &= \lambda_d \mathcal L_d + \lambda_c \mathcal L_c + \lambda_s \mathcal L_s + \lambda_{id1} \mathcal L_{id1} + \lambda_{id2} \mathcal L_{id2}
\end{align}
where the $\lambda$'s are weightings. We call this model D). We can train this using gradient based optimization, except the weights of the teacher are not updated by the gradient. The weights of the teacher are an exponential moving average of the students weights.

We also define model E) where we modify (1) to
\begin{align}
	\mathcal L_d &= \sum_{I \in \{(I_s)^g_1, (I_s)^g_2\}} \sum_{I' \in V - \{ I \}} H(D(SE_t(I), CE(I_c)), D(SE_s(I'), CE(I_c))).
\end{align}
Hoping that calculating the loss after the decoder will weigh the most relevant differences between the student and teacher higher. Relevant meaning making a greater difference to the final issues.

@@

\begin{center}
\includegraphics[width=14cm]{comparison}	
\end{center}

@@

\subsection{Quantitative Evaluation}
